### 2022-12-19 Draft3.Rmd
R\ : added backslash after R

---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-12-09"
abstract: >
  This R\ package translates R\ objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R\ functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R\ code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic representation, this new
  feature is a futher step towards reproducible science. User-defined hooks
  allow to extend the package by mapping new variable names and functions to
  new MathML and LaTeX entities. We provide a few working examples that
  illustrate the use of the package in a little scientific report and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/gondan-rochon/index.html.en
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/alfarone/index.html.en
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
preamble:
- \usepackage{longtable}
- \usepackage{cancel}
---

```{r setup, include=FALSE}
# To do
#
# paste(x, y, z)
# Spacing
# Comma separated lists
# Ellipsis
# Radicals
# Subsets
# Greek letters in MathJax
# Typeface
# Stylechanges
# Phantoms

library(mathml)
library(plotly)
library(ggplot2)
library(kableExtra)

mathout <- function(..., flags=NULL)
{
  if(knitr::is_html_output())
    return(mathml(..., flags=c(flags, list(cat=TRUE))))

  if(knitr::is_latex_output())
    return(mathjax(..., flags=c(flags, list(cat=TRUE))))

  warning("no output specified")  
  mathjax(..., flags=c(flags, list(cat=TRUE)))
}

catmath <- function(term, ...)
{
  s <- substitute(term)
  mathout(canonical(s), ...)
}
```

# Introduction

The R\ extension of the markdown language [@Xie2020] enables reproducible
statistical reports with nice typesetting in HTML, Microsoft Word, and Latex.
Since recently [@R, version 4.2], R\'s manual pages include support for
mathematical expressions [@Sarkar2022], which already is a big improvement.
However, rules for the translation of R\'s built-in language elements to their
mathematical representation are still lacking. So far, R\ expressions such
as `pbinom(k, N, p)` are printed as they are; pretty mathematical formulae
like $P_\mathrm{Bi}(X \le k; N, p)$ require explicit Latex commands, that
is, `P_\mathrm{Bi}\left(X \le k; N, p\right)`. Except for minimalistic examples,
these commands are tedious to type in and hard to read in their source code.

The present R\ package defines a set of rules for an _automatic_ translation of
R\ expressions to mathematical output in RMarkdown documents [@Xie2020] and
ShinyApp webpages [@Chang2022]. The translation is done by an embedded Prolog
interpreter that maps nested expressions recursively to MathML and
LaTeX/MathJax, respectively. User-defined hooks enable to extend the set of
rules, for example, to represent specific R\ elements by custom mathematical
signs.

The main feature of the package that the same R\ expressions and equations are
used for both mathematical typesetting and calculations. This saves time and
reduces mistakes, as well be illustrated below.

The paper is organized as follows. Section\ 2 presents the background of the
package, including Prolog and the R\ package \pkg{rolog} and the two main classes
of rules for translating R\ objects to mathematical expressions. Section\ 3
illustrates the main features of the \pkg{mathml} package, potential issues and
workarounds using examples from the day-to-day perspective of the user.
Sections\ 4 and\ 5 present two case studies for the use of the package in a
scientific report and a ShinyApp web page. Section\ 6 concludes with a
discussion and ideas for further development. A complete list of the R\ elements
covered by the package is found in the Appendix.

# Background

The translation of R\ expressions to mathematical output is achieved through a
Prolog interpreter. Prolog is a classical logic programming language with many
applications in expert systems, computer linguistics and symbolic artificial
intelligence. The main strength of Prolog is its concise representation of facts
and rules for the representation of knowledge and grammar, as well as its
efficient built-in search engine for closed world domains. As it is well-known,
R\ is a statistical programming language for data analysis and statistical
modeling which is widely used in academia and industry. Besides the core
library, a lot of packages have been developed for all kinds of statistical
problems, including statistics-based artificial intelligence tools such as
neural networks for machine learning and deep learning. Whereas Prolog is weak
in statistical computation, but strong in symbolic manipulation, the converse
may be said for the R\ language. The \pkg{rolog} package bridges this gap by
providing an interface to a SWI-Prolog distribution in R. The communication
between the two systems is mainly in the form of queries from R\ to Prolog, but
two predicates allow Prolog to ring back and evaluate terms in R.

For a first illustration of the package, we consider the binomial probability.

```{r}
term <- quote(pbinom(k, N, p))
term
```

The term is quoted to avoid its immediate evaluation which would raise an error
anyway since the variables `k`, `N`, `p` have not yet been defined. Experienced
readers will remember that the quoted expression is a short form for

```r
term <- call("pbinom", as.name("k"), as.name("N"), as.name("p"))
```

As is seen from the output above, the variable `term` is not assigned the result
of the calculation, but an R\ "call"
[see, e.g., @Wickham2019, for details on "non-standard evaluation"]. This call
can eventually be evaluated with `eval(...)`,

```{r}
k <- 10
N <- 22
p <- 0.4
eval(term)
```

The R\ package \pkg{mathml} can now be used to render the call in MathML, that
is the math dialect for HTML webpages or in MathJax/LaTeX, as shown
below (some of the curly braces are not really needed in this simple example,
but are necessary in edge cases).

```{r}
library(mathml)
mathjax(term)
```

We can include the output in a RMarkdown document by specifying `results='asis'`
in the R\code chunk, as is done in the next example. The
R\ function `mathout(...)` is just a wrapper that invokes `mathml(...)` for HTML
output and `mathjax(...)` for LaTeX output.

```{r, results='asis'}
mathout(term)
```

At the Prolog end, there is a predicate `math/2` that translates the
call `pbinom(K, N, Pi)` into a "function" `fn/2` with the name `P_Bi`, one
argument `X =< K`, and the two parameters `N` and `Pi`.

```prolog
math(pbinom(K, N, Pi), M)
 => M = fn(subscript('P', "Bi"), (['X' =< K] ; [N, Pi])).
```

Thus, the predicate `math/2` could be considered a "macro" that translates a
mathematical element (here, `pbinom(K, N, Pi)`) to a different mathematical
element, namely `fn(..., ...)`. The low-level predicates `ml/3` and `jax/3`
convert mathematical elements to MathML or LaTeX, respectively.

```prolog
ml(Flags, fn(Name, (Args ; Params)), M)
 => ml(Flags, Name, N),
    ml(Flags, paren(list(op(;), [list(op(','), Args), list(op(','), Params)])), X),
    M = mrow([N, mo(&(af)), X]).
```

The relevant rule for `ml/3` builds the MathML entity `mrow([N, mo(&(af)), X])`,
with `N` representing the function's name and `X` its arguments and parameters,
enclosed in parentheses. The flags can be used for context-sensitive
translation (see, e.g., the errors below).

# Package \pkg{mathml} in practice

\pkg{mathml} is an R\ package for pretty mathematical representation of
R\ functions and objects in data analysis, scientific reports and interactive
web content. The currently supported features are listed below, roughly
following the order proposed by [@murrell2000].

## Basic elements

\pkg{mathml} contains several basic elements used in everyday mathematical
expressions, such as numbers, Latin and Greek letters, multi-letter identifiers,
accents, subscripts, and superscripts.

```{r, results='asis'}
term <- quote(1 + 2L + "a" + a + abc + phi + Phi + varphi + hat(b)[i, j]^2L
              + NaN + NA + TRUE + FALSE + Inf + -Inf)
mathout(term)
```

todo: parentheses around -infty, see also $(no)^{p+q}$ below.

## Operators and parentheses

Basic arithmetic operators and parentheses are translated as they are, as
illustrated below.

```{r, results="asis"}
term <- quote(a - ((b + c)) - d*e + f*(g + h) + i/j + k^(l + m) + {n*o}^{p + q})
mathout(term)
```

For multiplications involving only numbers and symbols, the multiplication sign
is omitted. This heuristic does not always produce the desired result;
therefore, \pkg{mathml} defines the extra functions `dot/2`, `nodot/2`,
and `times/2` that produce the respective multiplication signs.

```{r, results="asis"}
term <- quote(dot(d, e) + nodot(f, g + h) + times(X, Y))
mathout(term)
```

Note that for the standard operators with known precedence, \pkg{mathml} is able
to detect if parentheses are needed; for example, a parenthesis is automatically
put around `g + h` in the `nodot`-example and around `n*o` above. However, we
also note an unecessary parenthesis around `l + m` above. This parenthesis
results is a consequence of `quote(a^(b + c))` actually producing a nested
R\ call of the form `'^'(a, (b + c))` instead of `'^'(a, b + c)`:

```{r}
term <- quote(a^(b + c))
paste(term)
```

This feature is unfortunate because the extra parenthesis around `b + c` is not
needed and yields to the extra parenthesis. The preferred result is obtained by
using the functional representation `quote('^'(k, l + m))` of the power, or, as
a workaround, curly braces (see last summand above).

## Custom operators

While the parentheses typically follow the precedence in standard infix
operators, undesirable results may be obtained in custom operators.

```{r, results='asis'}
term <- quote(mean(X) %+-% 1.96 * s / sqrt(N))
mathout(term)
```

As can be seen, \pkg{mathml} reminds the user that it is not possible to change
the precedence of custom operators in R\, and that expressions with such
operators are evaluated strictly from left to right. As in the
previous example, the solution is to either use the functional form of the
operator or a curly brace to enforce the correct operator precedence.

```{r, results='asis'}
term <- quote(`%+-%`(mean(X), 1.96 * s / sqrt(N)))
term <- quote(mean(X) %+-% {1.96 * s / sqrt(N)})   # the same
mathout(term)
```

More operators are shown in Table \@ref(tab:custom-operators).

```{r custom-operators, echo=FALSE}
op <- list(
  cross=quote(x %*% y),
  dot=quote(x %.% y),
  kronecker=quote(X %x% Y),
  'integer division'=quote(x %/% y),
  and=quote(A & B),
  vel=quote(A | B),
  xor=quote(xor(A, B)),
  not=quote(!A),
  equal=quote(X == Y),
  assign=quote(X <- Y),
  different=quote(x != y),
  tilde=quote(X ~ Y),
  approx=quote(x %~~% y),
  equiv=quote(x %==% y),
  congr=quote(x %=~% y),
  prop=quote(x %prop% y),
  'in'=quote(a %in% A),
  intersect=quote(intersect(A, B)),
  union=quote(union(A, B)),
  intersect=quote(is.null(intersect(A, B)))) # todo: check if duplicate

m <- lapply(op, FUN=mathjax, flags=list(cat=FALSE))

op <- as.character(op)
if(knitr::is_latex_output())
  op <- sapply(op, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op <- sapply(op, FUN=knitr:::escape_html)

t <- cbind(Operator=op, Output=m)
knitr::kable(t, caption="Custom operators in mathml", row.names=FALSE, escape=FALSE)
```

## R functions

The second example may be particularly of use when we deal with multiple conditions (see \@ref(case-study)). Here, for instance, we present a trivial one with different relation signs.

```{r, results='asis'}
f <- function(x, a)
{
    if(a == 0L) return( 0L );
    if(a <= 1L) return(-Inf);
    if(a > 1L) return(+Inf)
}

mathout(f)
```

Note that you may also assign a name to the function. The default name is in fact the greek letter lambda, but with `attributes()$name` the name costumization is performed smoothly.

```{r, results='asis'}
attributes(f)$name = quote(f)
mathout(f)
```

A few more elements are displayed in
Tables \@ref(tab:basic-elements) and \@ref(tab:arrows).

```{r basic-elements, echo=FALSE}
op1 <- list(
  lchoose=quote(lchoose(n, k)),
  sqrt=quote(sqrt(x)),
  mean=quote(mean(xy)),
  absolute=quote(abs(x))) # todo: check if duplicate
m1 <- lapply(op1, FUN=mathout, flags=list(cat=FALSE))

op1 <- as.character(op1)
if(knitr::is_latex_output())
  op1 <- sapply(op1, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op1 <- sapply(op1, FUN=knitr:::escape_html)

op2 <- list(
  twosidearrow=quote(x%<->%b),
  rightarrow=quote(x%->%y),
  leftarrow=quote(x%<-%y),
  uparrow=quote(x%up%y),
  down=quote(x%down%y),
  iff=quote(x%<=>%y),
  rArr=quote(x %=>% y),
  lArr=quote(x %<=% y),
  dblup=quote(x %dblup% y),
  dbldown=quote(x %dbldown% y))
m2 <- lapply(op2, FUN=mathout, flags=list(cat=FALSE))

op2 <- as.character(op2)
if(knitr::is_latex_output())
  op2 <- sapply(op2, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op2 <- sapply(op2, FUN=knitr:::escape_html)

t <- cbind(Operator=op1, Output=m1, Operator=op2, Output=m2)
knitr::kable(t, caption="Basic elements in mathml", row.names=FALSE, escape=FALSE)
```

## Ringing back to R

R\'s `integrate` function takes a number of arguments, the most important ones
being the function to integrate, and the lower and the upper bound of the
integration. For mathematical typesetting in the form
of $\int f(x)\, dx$, \pkg{mathml} needs to find out the name of the integration
variable. For that purpose, the underlying Prolog bridge provides a
predicate `r_eval/2` that calls R\ from Prolog. In the example below, this
predicate is used for a call of `formalArgs(args(sin))`, which returns the names
of the function arguments of `sin`, namely, `x`.

```{r, results='asis'}
term <- quote(integrate(sin, 0L, 2L*pi))
mathout(term)
```

## Names and order of arguments

One of R's great features is the possibility to refer to function arguments by
by their names, not only by their position in the list of arguments.
Unfortunately, the prolog-powered handler for `integrate/3` is rather rigid; it
accepts only three arguments in that particular order and without names, that
is, `integrate(lower=0L, upper=2L*pi, sin)`, would not print the desired result.

To "canonicalize" function calls with named arguments and arguments in unusual
order, \pkg{mathml} provides an auxilliary function `canonical(f, drop)` that
reorders the argument list of calls to known functions and, if `drop=TRUE`,
drops the function names.

```{r}
term <- quote(integrate(lower=0L, upper=2L*pi, sin))
canonical(term)
```

```{r, echo=TRUE, results='asis'}
# A custom function
g <- function(u)
{
  sin(u)
}

# Mixture of (partially) named and positional arguments in unusual order
term = call("integrate", lower=-Inf, upper=Inf, g)
attributes(term)$name = quote(g)
mathout(canonical(term))

# Even more complex
term <- call("*", 2L, call("$", call("integrate", low=-Inf, up=Inf, g), "value"))
attributes(term)$name = quote(g)
mathout(canonical(term))

# It is a bit of a mystery that R knows the result of this integral.
eval(term)
```

The extra R\ function `canonical()` applies `match.call()` to non-primitive R\
calls, basically cleaning up the arguments and bringing them into the correct
order.

## Customizing the package

In typical R\ functions, variable names are typically longer than just single
letters, which may yield unsatisfactory results in the mathematical output.

```{r, results='asis'}
term <- quote(pbinom(successes, Ntotal, prob))
mathout(term)
```

To improve the situation, \pkg{mathml} provides a simple hook into the prolog
predicate `math/2` described in the previous section.

```{r, results='asis'}
hook(quote(successes), quote(k))
hook(quote(Ntotal), quote(N))
hook(quote(prob), quote(pi))
mathout(term)
```

More complex hooks are possible. In the example below, `dnorm(X, M, S)` is
translated to the form `phi(X; M, S)` (instead of the default `varphi`).

```{r, results='asis'}
hook(call("dnorm", expression(X), expression(M), expression(S)), 
     call("fn", quote(phi),
       call(";", list(expression(X)), list(expression(M), expression(S)))))
term <- quote(dnorm(x, mu, sigma^2L))
mathout(term)
```

R's expressions are a bit misused to represent variables in the underlying
prolog engine.

## Operators

A list of possible operators is available here.

```{r, results='asis', echo=TRUE}
catmath(sum(x[i], i==1L, n))

# fixme: braces not needed when right to *
catmath(a * sum(b[i]))

# Multiplication (capital Pi) is missing
catmath(a * prod(b[i]))

catmath(integrate(sin, 0L, 2L*pi))

catmath(A & B)

catmath(A | B)

catmath(!A)

catmath(xor(A, B))

catmath(exp(2L*pi*i))

catmath(expm1(2L*pi*i))

catmath(log(x))

catmath(log10(x))

catmath(log2(x))

catmath(logb(x, e))

catmath(log1p(x))

catmath(ceiling(pi))

catmath(floor(pi))

catmath(sin(pi/2L))

catmath(cos(pi/2L))

catmath(tan(pi/2L))
```

## Grouping

At first glance, grouping expressions may seem trivial enough. However, when you want to deal with, for instance, exponentiation or multiplication, things can get more complicated. Let us consider a trivial example. If we want to calculate and evaluate the expression `a * (b + c)`, we find that R\ automatically gives precedence to the addition even if we do not specify it in the text, as in the following.

```{r, results='asis', echo=TRUE}
first <- quote(a)
second <- quote(b + c)
mathout(call("*", first, second))

# precedence
quote(a * (b + c))[[3]][[1]]
```

With exponentiation the situation can be less trivial. If we consider `x` to the power of `y + z`, following R\ notation we should add a parenthesis around the two terms. Otherwise R\ would consider only `y` as power of `x`.

```{r, results='asis', echo=TRUE}
catmath(x^y + z)

# mathout(call("+", call("^", as.symbol("x"), as.symbol("y")), as.symbol("z")))

catmath(x^(y + z))
```

Unfortunately, R\ automatically put `y + z` in parentheses even if, from a mathematical point of view, the brackets are not needed. To solve this issue two solutions are possible, the first one is cumbersome since it requires many specifications. The second, that can be considered a shortcut, uses a syntax similar to LateX and puts the power in curly brackets. In both cases the output produced is consistent with the mathematical notation.

```{r, results='asis', echo=TRUE}
# without parenthesis
mathout(canonical(call("^", as.symbol("x"), call("+", as.symbol("y"), as.symbol("z")))))

# shortcut
catmath(x^{y + z})
```

## Matrix and Vectors

Since also Matrices and Vectors are important elements in R\, we decided to implement them and render them graphically in the following ways.

```{r, results='asis', echo=TRUE}
vec <- 1:3
mathout(canonical(vec))
mathout(vec)

mat <- matrix(data = c(2L, 3L, 4L, 5L), nrow = 2, ncol = 2)
mathout(canonical(mat))
```

## Abbreviations

\pkg{mathml} allows also for the use of abbreviations in formulas. Let us consider here the t-statistic for independent samples with equal variance assumed. 
The procedure consists of three steps. The first one is to define the hooks to nicely print the variables. Second, we write the t-statistic specifying in the formula how to calculate the pooled variance with the function  `denote(abbr, expr, info)`.  Finally, since \pkg{mathml} also allows computation, we can directly compute the value of the t statistic with `eval()`.

```{r, results='asis', echo=TRUE}

hook(quote(x_1), quote(mean(x["1"])))
hook(quote(x_2), quote(mean(x["2"])))
hook(quote(n_1), quote(n["1"]))
hook(quote(n_2), quote(n["2"]))
hook(quote(s_1), quote(s["1"]))
hook(quote(s_2), quote(s["2"]))
hook(quote(s_p), quote(s["pool"]))

t <- quote(dfrac(x_1 - x_2, sqrt(denote(s_p^2L, frac((n_1 - 1L)*s_1^2L + (n_2 - 1L)*s_2^2L, n_1+n_2-2L), "the pooled variance.") * (dfrac(1L, n_1)+dfrac(1L,n_2)))))

mathout(t)

# An example

x_1 = 20.5
x_2 = 35.7
n_1 = 150
n_2 = 160
s_1 = 3.4
s_2 = 4.2

eval(t)
```

# A case study {#case-study}

For a more clear representation of what the \pkg{mathml} package is supposed to do, we provide here a case study, based on the work by Schwarz [-@schwarz1994], in which the features of \pkg{mathml} are implemented.

The reader may recognize that Schwarz' [-@schwarz1994] study can be traced back to the strand of decision-making models, where decision-making is assumed to be a process of
noisy accumulation of information over time [e.g., @ratcliff2016]. The aim of
Schwarz' [-@schwarz1994] study is to present a new explanation of redundancy gains when
observers make speeded responses to stimuli of different sources, and the same
information is presented on two channels. Schwarz [-@schwarz1994] describes a model that assumes
the superposition of channel-specific diffusion processes that eventually reach
an absorbing barrier to elicit the response. For a detailed description the
reader may refer to the original article.

### The model

Schwarz' [-@schwarz1994] model refers to two stimuli A and B, presented either
alone or in combination (redundant stimuli), with the redundant stimuli being
presented either simultaneously or with onset asynchrony $\tau$. The channel
activation is described as a two-dimensional Wiener process with
drifts $\mu_\mathrm A$, $\mu_\mathrm B$,
variances $\sigma^2_\mathrm A$, $\sigma^2_\mathrm B$,
correlation $\rho_\mathrm{AB}$ and initial
conditions $X_i(t=0)=0, i=\mathrm{A, B}$. The response is elicited when the
process reaches the absorbing barrier $c > 0$ for the first time. In combined
stimuli ("redundant" stimuli), the overall diffusion process
is $X_\mathrm{AB}(t) = X_\mathrm A(t) + X_\mathrm B(t)$ , which is again a
Wiener process with drift $\mu_\mathrm A + \mu_\mathrm B$ and
variance $\sigma^2_\mathrm A + \sigma^2_\mathrm B
  + 2\rho_\mathrm{AB}\sigma_\mathrm A\sigma_\mathrm B$.

In single-target trials, the first passages at $c$ are expected at
$E\left[D_i\right] = \frac{c}{\mu_i}, i = \mathrm{A, B}$, whereas in synchronous
redundant-target trials, we have
$E[D_\mathrm{AB}]= \frac{c}{\mu_\mathrm A + \mu_\mathrm B}$.

For asynchronous stimuli, Schwarz [-@schwarz1994] derived the expected
first-passage time $E[D_{\mathrm A(\tau)\mathrm B}]$ as a function of the
stimulus onset asyncrony $\tau > 0$ (Eq.\ 10),

```{r, results="asis", echo=TRUE}
f <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) *
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*c, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

mathout(f)
```

The observable response time is assumed to be the sum of $D$, the time
employed to reach the threshold for the decision, and $M$, denoting the of the
other processes.

$E[T]= E[D + M] = E[D] + E[M] = E[D] + \mu_M$

### Methods

Schwarz [-@schwarz1994] applied the model to Miller’s [-@miller1986] data from
a redundant signals task with 13 onset
asynchronies $0, \pm33, \pm67, \pm100, \pm133, \pm167, \pm\infty$ ms,
where 0 refers to synchronous AB, $\infty$ refers to the single-target
presentation, and negative $\tau$ denote those conditions in which B is
presented before A. Each condition was replicated 400 times. The 13 mean RTs and
standard deviations are given in Table 2 below.

Then, from Schwarz' [-@schwarz1994] Equations [2] [3] and [10], we calculated the expected first time passage for the respective $\tau$ and the parameters $\mu_ \mathrm A = 0.53, \mu_ \mathrm B = 1.32, \sigma_ \mathrm A = 4.43, \sigma_ \mathrm B = 18.3, c = 100, \mu_\mathrm M = 161$

````{r, echo=FALSE}
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = c(-Inf, -167, -133, -100, -67, -33, 0, 33, 67, 100, 133, 167, Inf)
c = 100
mu_M = 161
````

```{r, results="asis", echo=FALSE}
# Conditions

g <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{
    if( tau == Inf) return(dfrac(c, mu_A));
    if( tau == -Inf) return(dfrac(c, mu_B));
    if( tau > 0L)  return(f(tau, c, mu_A, sigma_A, mu_B, sigma_B));
    if( tau == 0L) return(dfrac(c, mu_A + mu_B));
    if( tau < 0L) return(f(-tau, c,  mu_B, sigma_B, mu_A, sigma_A))
}

mathout(g)
```

For instance, we can see that with $\tau = -167$ the expected first passage time is:

```{r, echo=FALSE}
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = -167
c = 100
mu_M = 161

g(tau, c, mu_A, sigma_A, mu_B, sigma_B)
```

To which we have to add $\mu_M = 161$ (which represents the amount of time that the movement of providing the answer takes) and we obtain:

```{r, echo=FALSE}
g(tau, c, mu_A, sigma_A, mu_B, sigma_B) + mu_M
```

Similarly, we obtain the first-time passage for every $\tau$:

```{r, results="asis", echo=FALSE}

G <- Vectorize(g, vectorize.args = 'tau')

h <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
{
  G(tau, c, mu_A, sigma_A, mu_B, sigma_B) + mu_M
}

mathout(h)
```

```{r, echo=FALSE}

mu_M = 161
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = c(-Inf, -167, -133, -100, -67, -33, 0, 33, 67, 100, 133, 167, Inf)
c = 100

h(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
```

### Model fitting and Parameters Estimation

Then we proceeded with the model fitting, using the data provided by Miller [-@miller1986] study in Schwarz [-@schwarz1994]. First, we calculated the predicted means RT as $E[RT]=E[T]+\mu_M$.

We briefly remind that in diffusion models there is across-trial variability, and the drift rates are assumed to be normally distributed [@ratcliff2008] with at the first passage of $c$ $X(t) \sim N(\mu t , \sigma^2 t)$. Where $T$ is expressed as $T= \arg \min_t X(t)\geq c$ with $T \sim IG (c, \mu, \sigma^2)$ with mean $E(T)=\displaystyle \frac{c}{\mu}$ and variance $Var(T)=\displaystyle \frac{c\sigma^2}{\mu^3}$. Furthermore, for the Central Limit Theorem $\bar{T} = \displaystyle \frac{\Sigma_iT_i}{N} \dot\sim N (\displaystyle \frac{c}{\mu} , \displaystyle \frac{c\sigma^2}{\mu^3})$

For these reasons, in order to obtain the Goodness of Fit measure we proceeded with a z-standardisation.

```{r, results="asis", echo=FALSE}
zf <-  function(M, s, N, tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
{
  E <- h(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
  dfrac(M-E, s/sqrt(N))
}

mathout(zf)
```

```{r, echo=FALSE}
# The vector X contains the observed values, the vector s their observed standard deviation and N = 400 are the number of trials.

M = c(231, 234, 230, 227, 228, 221, 217, 238, 263, 277, 298, 316, 348)
s = c(56, 58, 40, 40, 32, 28, 28, 28, 26, 30, 32, 34, 92)
N = c(400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400)

zn <- zf(M, s, N, tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
zn
```

Then, we calculated the Goodness of Fit as follows:

$X^2=\Sigma_\tau z^2_\tau \sim\chi^2(13-5\  dgf)$.

```{r, results="asis", echo=FALSE}
par <- c(mu_A = 0.53, sigma_A = 4.4, mu_B = 1.32, sigma_B = 18.3, mu_M = 161)

# Goodness of Fit

gof <- function(par, tau, M, s, N)
{
  zn <- zf(M, s, N, tau, c=100L, mu_A=par["mu_A"], sigma_A=par["sigma_A"], mu_B=par["mu_B"], sigma_B=par["sigma_B"], mu_M=par["mu_M"])
  sum(zn^2L)
}

mathout(gof)
```

```{r, results="asis", echo=FALSE}
n <- gof(par, tau, M, s, N)
n
pchisq(n, 8, lower.tail = FALSE, log.p = FALSE)
```

```{r, results="asis", echo=FALSE}
lsq = function(tau, M, s, N)
{
  start = c(mu_A = 0.3, sigma_A = 2, mu_B = 1 , sigma_B = 5, mu_M = 200)
  optim(start, fn = gof,
    tau = tau, M = M, s = s, N = N, hessian = TRUE, control = list(fnscale = 1))
}

mathout(lsq)

# Parameters' estimates
fit <-  lsq(tau, M, s, N)

est = fit$par
se = sqrt(diag(solve(fit$hessian)))
cilo = qnorm(0.025, est, se)
ciup = qnorm(0.975, est, se)

# Lower bound
#cilo

# Upper bound
#ciup

```

### Results

The calculated Goodness of Fit value is $X^2= 30.54$, $p= 0.0002$. Additionally, we proceeded with the parameters' estimation with the Least Squares Method.

The values of the estimated parameters reported in the table below are consistent with those reported by Schwarz [-@schwarz1994].

```{r}
# Table 1: Estimated Model's Parameters
#
# | Parameters | Estimate | 95% Confidence Interval
# | :---: | :---: | :---:|
# | $ \mu_ \mathrm A $| 0.53 | 0.51, 0.56
# | $ \sigma_ \mathrm A $| 4.4 | 3.8, 5.1
# | $ \mu_ \mathrm B $| 1.32 | 1.2, 1.4
# | $ \sigma_ \mathrm B $| 18.3 | 11.6, 24.9
# | $ \mu_ \mathrm M $| 161 | 157, 165
```

Furthermore, we can observe from the values reported in the Table below that, with few exceptions (e.i. $\tau= 100$) Miller's [-@miller1986] observed values are quite similar to the one predicted by Schwarz' [-@schwarz1994] model. Showing that, despite the calculated Goodness of Fit measure, the model proposed by Schwarz [-@schwarz1994] sufficiently fit the data.

```{r}
# : Table 2: Predicted Means RT and Observed Means RT and SDs from Miller's [-@miller1986] study in Schwarz [-@schwarz1994]
#
# | SOA      | Predicted Mean RT | Observed Mean RT | Observed SD |
# |:--------:|:-----------------:|:----------------:|:-----------:|
# $-\infty$| 237 | 231 | 56 |
# $-167$   | 231 | 234 | 58 |
# $-133$   | 230 | 230 | 40 |
# $-100$   | 228 | 227 | 40 |
# $-67$    | 226 | 228 | 32 |
# $-33$    | 222 | 221 | 28 |
# $0$      | 215 | 217 | 28 |
# $33$     | 239 | 238 | 28 |
# $67$     | 262 | 263 | 26 |
# $100$    | 283 | 277 | 30 |
# $133$    | 300 | 298 | 32 |
# $167$    | 313 | 316 | 34 |
# $\infty$ | 350 | 348 | 92 |
#
# _Note_.  Parameters: $\mu_ \mathrm A = 0.53, \mu_ \mathrm B = 1.32, \sigma_ \mathrm A = 4.4, \sigma_ \mathrm B = 18.3, c = 100, \mu_\mathrm M = 161$
```

## Mistakes

The case study presented above displays several possible ways of how \pkg{mathml} can be a useful tool in data representation and calculation. Furthermore, the careful reader might have also noticed that \pkg{mathml} allows also for a more smoothly detection of possible mistakes in writing.

An interesting use case of the package is in this sense the readability of self-written
functions in R\ code. R\ code is notoriously hard to read, as are most programming
languages [@green1977]. The poor legibility of the language is one of the main
sources of mistakes. For illustration, we consider Equation 10 in
Schwarz [-@schwarz1994], namely the same equation reported above. The first version has a wrong parenthesis, which is barely visible in the R code.

````{r, results="asis", echo=TRUE}
f <- function(tau)
{ dfrac(c, mu["A"]) + (dfrac(1L, mu["A"]) - dfrac(1L, mu["A"] + mu["B"]) * 
    ((mu["A"]*tau - c) * pnorm(dfrac(c - mu["A"]*tau, sqrt(sigma["A"]^2L*tau)))
      - (mu["A"]*tau + c) * exp(dfrac(2L*mu["A"]*tau, sigma["A"]^2L))
        * pnorm(dfrac(-c - mu["A"]*tau, sqrt(sigma["A"]^2L*tau)))))
}

attributes(f)$name = quote(f)
mathout(f)
````

But, if the function is rendered in MathML, the wrong parenthesis appears as a curly
brace, and we immediately see the mistake. The correct version is shown below
for comparison.

````{r, results="asis", echo=TRUE}
f <- function(tau)
{ dfrac(c, mu["A"]) + (dfrac(1L, mu["A"]) - dfrac(1L, mu["A"] + mu["B"])) * 
    ((mu["A"]*tau - c) * pnorm(dfrac(c - mu["A"]*tau, sqrt(sigma["A"]^2L*tau)))
      - (mu["A"]*tau + c) * exp(dfrac(2L*mu["A"]*tau, sigma["A"]^2L))
        * pnorm(dfrac(-c - mu["A"]*tau, sqrt(sigma["A"]^2L*tau))))
}

attributes(f)$name = quote(f)
mathout(f)
````

This example may seem trivial. However, as the reader well knows, a missed parenthesis is often the cause of untraceable errors and contradictory results. Thus, rendering the expression in MathMl improves the readability of the text, allowing better identification of the sources of possible errors.

## Errors

An additional feature that this package offers is related to its application in educational scenarios. The functions here reported are intended to identify and highlight possible errors that students make, as well as provide the correct answer for the task.

In this first example the function highlights that the student forgot to add the left-hand side of the expression and it does that by striking the missing part. In this case the letter A and the operator `+`.
Similarly the function `omit_right` does the same for the omissions in the right-hand side of the expression. Finally the function `omit()` can be used, for instance, to highlight the omission in the elements of a list.

```{r, results="asis"}
t <- quote(omit_left(a + b))
mathout(t)
```

```{r, results="asis"}
t <- quote(omit_right(a + b))
mathout(t)
```

```{r, results="asis"}
t <- list(quote(a), quote(b), quote(omit(c)))
mathout(t)
```

```{r omit, echo=FALSE}
op <- list(left=quote(omit_left(a + b)), right=quote(omit_right(a + b)), list=quote(omit(c)))
m <- lapply(op, FUN=mathout, flags=list(cat=FALSE))

op <- as.character(op)
if(knitr::is_latex_output())
  op <- sapply(op, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op <- sapply(op, FUN=knitr:::escape_html)

t <- cbind(Operator=op, Output=m)
knitr::kable(t, caption="Omit", row.names=FALSE, escape=FALSE)
```

The functions `add_left()`, `add_right()`, `add()` reported below follow the same way of reasoning, but they account for the mistakes made in adding incorrect elements in the expression. They highlight the redundant element placing it in a box.

```{r, results="asis"}
t <- quote(add_left(a + b))
mathout(t)
```

```{r, results="asis"}
t <- quote(add_right(a + b))
mathout(t)
```

```{r, results="asis"}
t <- list(quote(a), quote(b), quote(add(c)), quote(d))
mathout(t)
```

```{r add, echo=FALSE}
op <- list(left=quote(add_left(a + b)), right=quote(add_right(a + b)))
m <- lapply(op, FUN=mathout, flags=list(cat=FALSE))

op <- as.character(op)
if(knitr::is_latex_output())
  op <- sapply(op, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op <- sapply(op, FUN=knitr:::escape_html)
t <- cbind(Operator=op, Output=m)

knitr::kable(t, caption="Add", row.names=FALSE, escape=FALSE)
```

Finally the function `instead( , )` place a curly bracket under the incorrect element  and provides its correct version under the curly bracket. In order to provide the correction with this function one must fill it with its wrong and correct terms in this order: `instead(wrong, correct)`. An example of its use its provided below.

```{r, results="asis"}
t <- quote(instead(a, b) + c)
mathout(t)
```

It is also possible to use flags for context-dependent rendering. The `highlight` flag is the default one. It shows where the error is and what is its solution. The `ignore` flag allows the program to ignore the fact that there is an error and print the expression as if it were correct. The third flag, `fix`, tells you what is the term that you are supposed to fix without giving the solution. Finally, the last flag `asis` returns the incorrect expression as the student wrote it, without providing any solutions or corrections.  Below is provided the code with its comment and an example of use.

Let us consider for example a student that instead of writing `b + c` wrote `a+c`.

With the `asis` flag we report the incorrect expression provided by the student:

```{r, results="asis"}
# It prints the incorrect version
t <- quote(instead(a, b) + c)
mathout(t, flags=quote(error(asis)))
```

Then, we may want to underline the incorrect term and provide the correct solution.

```{r, results="asis"}
# It highlights what has been mistakenly inserted
mathout(t, flags=quote(error(highlight))) # default
```

Alternatively, we can put the correct solution in a box to let the student focus on its correction.

```{r, results="asis"}
# It highlights what has been corrected
mathout(t, flags=quote(error(fix)))
```

Finally, we can simply provide the correct solution to the problem wit the flag `ignore`.

```{r, results="asis"}
# It prints the correct version
mathout(t, flags=quote(error(ignore)))
```

# Conclusions

This package expands the possibilities of R\ by allowing calculations and graphical rendition at the same time. This makes it easier for the user to notice writing errors and correct them smoothly. Building on current features of R\ and existing packages for displaying formulas in R\ (e.g. Murrell and Ihaka [-@murrell2000], Allaire et al. [-@allaire2018]), the \pkg{mathml} package bridges the gap between computational needs, presentation of results, and their reproducibility. To do so, it uses the Prolog's grammar rules and translates the R\ expressions to MathML. Then, two possible solutions for typesetting are available to export R\ Markdown documents in different formats or ShinyApp webpages: MathML and MathJax.

The article and Appendix show examples of how the package can be used and what operations are currently possible. Additionally, \pkg{mathml} allows for user's custom hooks to provide a better definition of the needed functions.

In its current version \pkg{mathml} has some limitations. For example, it is still not possible to use the functions of \pkg{mathml} for writing intratextual formulas, while it is still more convenient to adopt the usual LateX notation. However, we do not consider this to be a serious limitation, since the most productive use of \pkg{mathml} still comes from more complicated calculations. Nevertheless, in the future versions of this package we will address this issue, to broaden its spectrum of possible applications.

\pkg{mathml} is available for R\ version (…) and later, and can be easily installed using the usual `install.packages("mathml")`. The source code of the package is found at (…) including installation instructions for Unix, Windows and macOS.

# Acknowledgments