### 2023-01-30 rj/mathml.Rmd
initial checks

---
title: "`mathml`: Translate R Expressions to MathML and LaTeX"
date: "2023-01-28"
abstract: >
  This R\ package translates R\ objects to suitable elements in MathML or LaTeX,
  thereby allowing for a pretty mathematical representation of R\ objects and
  functions in data analyses, scientific reports and interactive web content. In
  the RMarkdown document rendering language, R\ code and mathematical content
  already exist side-by-side. The present package allows to use the same
  R\ objects for both data analysis and typesetting in documents or web content.
  This tightens the link between the statistical analysis and its verbal
  description or symbolic representation, which is another step towards
  reproducible science. User-defined hooks allow to extend the package by
  mapping specific variables or functions to new MathML and LaTeX entities. A
  number of examples and a case study illustrate the use of the package in a
  scientific report.
draft: true
author:  
  - name: Matthias Gondan
    affiliation: Universität Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/gondan-rochon/index.html.en
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/alfarone/index.html.en
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: Universität Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: bibliography.bib
preamble:
- \usepackage{cancel}
- \usepackage{amsmath}
Vignette: >
  %\VignetteIndexEntry{mathml: Translate R expressions to MathML and LaTeX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(mathml)

mathout <- function(..., flags=NULL, env=parent.frame())
{
  if(knitr::is_html_output())
    return(mathml(..., flags=c(flags, list(cat=TRUE)), env=env))

  if(knitr::is_latex_output())
    return(mathjax(..., flags=c(flags, list(cat=TRUE)), env=env))

  warning("mathout: no knitr output specified")  
  mathjax(..., flags=c(flags, list(cat=TRUE)), env=env)
}

catmath <- function(term, ...)
{
  s <- substitute(term)
  mathout(canonical(s), ...)
}
```

# Introduction

The R\ extension of the markdown language [@Xie2020] enables reproducible
statistical reports with nice typesetting in HTML, Microsoft Word, and Latex.
Since recently [@R, version 4.2], R\'s manual pages include support for
mathematical expressions [@Sarkar2022;@Viechtbauer2022], which already is a big
improvement. However, rules for the mapping of built-in language elements to
their mathematical representation are still lacking. So far, R\ expressions such
as `pbinom(k, N, p)` are printed as they are; pretty mathematical formulae
such as \(P_{\mathrm{Bi}}(X \le k; N, p)\) require explicit Latex commands, that
is, `P_{\mathrm{Bi}}\left(X \le k; N, p\right)`. Except for minimalistic use
cases, these commands are tedious to type in and their source code is hard to
read.

The present R\ package defines a set of rules for the automatic translation of
R\ expressions to mathematical output in RMarkdown documents [@Xie2020] and
ShinyApp webpages [@Chang2022]. The translation is done by an embedded Prolog
interpreter that maps nested expressions recursively to MathML and
LaTeX/MathJax, respectively. User-defined hooks enable to extend the set of
rules, for example, to represent specific R\ elements by custom mathematical
signs.

The main feature of the package is that the same R\ expressions and equations
can be used for both mathematical typesetting and calculations. This saves time
and reduces mistakes, as will be illustrated below.

The paper is organized as follows. Section\ 2 presents the technical background
of the package, including the two main classes of Prolog rules for translating
R\ objects to mathematical expressions. Section\ 3 illustrates the main features
of the \CRANpkg{mathml} package, potential issues and workarounds using examples
from the day-to-day perspective of a user. Section\ 4 present a case study for
the use of the package in a scientific report. Section\ 5 concludes with a
discussion and ideas for further development.

# Background

The translation of R\ expressions to mathematical output is achieved through a
Prolog interpreter from R\ package \CRANpkg{rolog} [@rolog]. Prolog is a classical logic
programming language with many applications in expert systems, computer
linguistics and symbolic artificial intelligence. The main strength of Prolog is
its concise representation of facts and rules for knowledge and grammar, as well
as its efficient built-in search engine for closed world domains. As it is
well-known, R\ is a statistical programming language for data analysis and
statistical modeling. Whereas Prolog is weak in statistical computation, but
strong in symbolic manipulation, the converse may be said for the R\ language.
The \pkg{rolog} package bridges this gap by providing an interface to a SWI-Prolog
distribution [@swipl] in R. The communication between the two systems is mainly
in the form of queries from R\ to Prolog, but two predicates allow Prolog to
ring back and evaluate terms in R.

For a first illustration of the \pkg{mathml} package, we consider the binomial
probability.

```{r}
term <- quote(pbinom(k, N, p))
term
```

The term is quoted to avoid its immediate evaluation (which would raise an error
anyway since the variables `k`, `N`, `p` have not yet been defined). Experienced
readers will remember that the quoted expression above is a short form for

```r
term <- call("pbinom", as.name("k"), as.name("N"), as.name("p"))
```

As is seen from the output above, the variable `term` is not assigned the result
of the calculation, but an R\ call
[see, e.g., @Wickham2019, for details on "non-standard evaluation"]. Such a call
can eventually be evaluated with `eval()`,

```{r}
k <- 10
N <- 22
p <- 0.4
eval(term)
```

The R\ package \pkg{mathml} can now be used to render the call in MathML, that
is the dialect for mathematical elements on HTML webpages or in MathJax/LaTeX,
as shown below (some of the curly braces are not really needed in this simple
example, but are necessary in edge cases).

```{r}
library(mathml)
mathjax(term)
```

We can include the output in a RMarkdown document by specifying `results='asis'`
in the R\ code chunk, as is done in the next example. The
R\ function `mathout()` has been defined in this vignette; it invokes `mathml()`
for HTML output and `mathjax()` for LaTeX output.

```{r, results='asis'}
mathout(term)
```

At the Prolog end, a predicate `math/2` translates the call `pbinom(K, N, Pi)`
into a "function" `fn/2` with the name `P_Bi`, one argument `X =< K`, and the
two parameters `N` and `Pi`.

```prolog
math(pbinom(K, N, Pi), M)
 => M = fn(subscript('P', "Bi"), (['X' =< K] ; [N, Pi])).
```

Thus, the predicate `math/2` could be considered a "macro" that translates a
mathematical element (here, `pbinom(K, N, Pi)`) to a different mathematical
element, namely `fn(Name, (Args ; Pars))`. The low-level predicate `ml/3` is
used to convert these basic elements to MathML.

```prolog
ml(Flags, fn(Name, (Args ; Pars)), M)
 => ml(Flags, Name, N),
    ml(Flags, paren(list(op(;), [list(op(','), Args), list(op(','), Pars)])), X),
    M = mrow([N, mo(&(af)), X]).
```

The relevant rule for `ml/3` builds the MathML entity `mrow([N, mo(&(af)), X])`,
with `N` representing the name of the function and `X` its arguments and
parameters, enclosed in parentheses. A corresponding rule `jax/3` does the same
for MathJax/LaTeX. A list of flags can be used for context-sensitive
translation (see, e.g., the section on errors below).

# Package \pkg{mathml} in practice

\pkg{mathml} is an R\ package for pretty mathematical representation of
R\ functions and objects in data analysis, scientific reports and interactive
web content. The currently supported features are listed below, roughly
following the order proposed by [@murrell2000].

## Basic elements

\pkg{mathml} handles the basic elements of everyday mathematical expressions,
such as numbers, Latin and Greek letters, multi-letter identifiers, accents,
subscripts, and superscripts.

```{r, results='asis'}
term <- quote(1 + -2L + a + abc + "a" + phi + Phi + varphi + roof(b)[i, j]^2L)
mathout(term)

term <- quote(round(3.1415, 3L) + NaN + NA + TRUE + FALSE + Inf + (-Inf))
mathout(term)
```

An expression such as `1 + -2` may be considered unsatisfactory from an
aesthetical perspective. It is correct R\ syntax, though, and is reproduced
accordingly, without the parentheses. Parentheses around negated numbers or
symbols can be added as shown above for `+ (-Inf)`.

To avoid name clashes with package `stats`, `roof()` is used to put a hat
on a symbol (see next section for further decorations). Note that an 
R\ function `roof()` does not exist in base R, it is provided by the package
for convenience and points to the identity function.

## Decorations

The package offers some support for different fonts as well as accents and
boxes etc. Internally, these decorations are implemented as identity functions,
so they can be introduced into R expressions without side-effects.

```{r, results='asis'}
term <- quote(bold(b[x, 5L]) + bold(b[italic(x)]) + italic(ab) + italic(42L))
mathout(term)

term <- quote(tilde(a) + mean(X) + box(c) + cancel(d) + phantom(e) + prime(f))
mathout(term)
```

Note that the font styles only affect the display of identifiers, whereas
numbers, character strings etc. are left untouched.

## Operators and parentheses

Arithmetic operators and parentheses are translated as they are, as illustrated
below.

```{r, results="asis"}
term <- quote(a - ((b + c)) - d*e + f*(g + h) + i/j + k^(l + m) + (n*o)^{p + q})
mathout(term)

term <- quote(dot(a, b) + frac(1L, nodot(c, d + e)) + dfrac(1L, times(g, h)))
mathout(term)
```

For multiplications involving only numbers and symbols, the multiplication sign
is omitted. This heuristic does not always produce the desired result;
therefore, \pkg{mathml} defines alternative R\ functions `dot()`, `nodot()`,
and `times()`. These functions calculate a product and produce the respective
multiplication signs. Similarly, `frac()` and `dfrac()` can be used for small
and large fractions.

For standard operators with known precedence, \pkg{mathml} is generally able to
detect if parentheses are needed; for example, parentheses are automatically
placed around `g + h` in the `nodot`-example. However, we note unnecessary
parentheses around `l + m` above. Thes parentheses are a consequence
of `quote(a^(b + c))` actually producing a nested R\ call of the
form `'^'(a, (b + c))` instead of `'^'(a, b + c)`:

```{r}
term <- quote(a^(b + c))
paste(term)
```

For the present purpose, this feature is unfortunate because extra parentheses
around `b + c` are not needed. The preferred result is obtained by using the
functional form `quote('^'(k, l + m))` of the power, or curly braces as a
workaround (see `p + q` above).

## Custom operators

Whereas in standard infix operators, the parentheses typically follow the rules
for precedence, undesirable results may be obtained in custom operators.

```{r, results='asis'}
term <- quote(mean(X) %+-% 1.96 * s / sqrt(N))
mathout(term)
term <- quote('%+-%'(mean(X), 1.96 * s / sqrt(N))) # functional form of '%+-%'
term <- quote(mean(X) %+-% {1.96 * s / sqrt(N)})   # the same
mathout(term)
```

The example is a reminder that it is not possible to define the precedence of
custom operators in R, and that expressions with such operators are evaluated
strictly from left to right. Again, the problem can be worked around by the
functional form of the operator, or a curly brace to hide the parenthesis but
enforce the correct operator precedence.

More operators are shown in Table \@ref(tab:custom-operators), including the
suggestions by Murrell and Ihaka [-@murrell2000] for graphical annotations and
arrows in R\ figures.

```{r custom-operators, echo=FALSE}
op1 <- list(
  "A\\ %*%\\ B"=quote(A %*% B),
  "A\\ %.%\\ B"=quote(A %.% B),
  "A\\ %x%\\ B"=quote(A %x% B),
  "A\\ %/%\\ B"=quote(A %/% B),
  "A\\ %%\\ B"=quote(A %% B),
  "A\\ &\\ B"=quote(A & B),
  "A\\ |\\ B"=quote(A | B),
  "xor(A,\\ B)"=quote(xor(A, B)),
  "!A"=quote(!A),
  "A\\ ==\\ B"=quote(A == B),
  "A\\ <-\\ B"=quote(A <- B))

m1 <- lapply(op1, FUN=mathout, flags=list(cat=FALSE))

op1 <- names(op1)
if(knitr::is_latex_output())
  op1 <- sapply(op1, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op1 <- sapply(op1, FUN=knitr:::escape_html)

op2 <- list(
  "A\\ !=\\ B"=quote(A != B),
  "A\\ ~ B"=quote(A ~ B),
  "A\\ %~~%\\ B"=quote(A %~~% B),
  "A\\ %==%\\ B"=quote(A %==% B),
  "A\\ %=~%\\ B"=quote(A %=~% B),
  "A\\ %prop%\\ B"=quote(A %prop% B),
  "A\\ %in%\\ B"=quote(A %in% B),
  "intersect(A,\\ B)"=quote(intersect(A, B)),
  "union(A,\\ B)"=quote(union(A, B)),
  "crossprod(A,\\ B)"=quote(crossprod(A, B)),
  "is.null(A)"=quote(is.null(A)))

m2 <- lapply(op2, FUN=mathout, flags=list(cat=FALSE))

op2 <- names(op2)
if(knitr::is_latex_output())
  op2 <- sapply(op2, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op2 <- sapply(op2, FUN=knitr:::escape_html)

op3 <- list(
  "A\\ %<->%\\ B"=quote(A %<->% B),
  "A\\ %->%\\ B"=quote(A %->% B),
  "A\\ %<-%\\ B"=quote(A %<-% B),
  "A\\ %up%\\ B"=quote(A %up% B),
  "A\\ %down%\\ B"=quote(A %down% B),
  "A\\ %<=>%\\ B"=quote(A %<=>% B),
  "A\\ %=>%\\ B"=quote(A %=>% B),
  "A\\ %<=%\\ B"=quote(A %<=% B),
  "A\\ %dblup%\\ B"=quote(A %dblup% B),
  "A\\ %dbldown%\\ B"=quote(A %dbldown% B),
  " "="")

m3 <- lapply(op3, FUN=mathout, flags=list(cat=FALSE))

op3 <- names(op3)
if(knitr::is_latex_output())
  op3 <- sapply(op3, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op3 <- sapply(op3, FUN=knitr:::escape_html)

t <- cbind(Operator=op1, Output=m1,
  Operator=op2, Output=m2,
  Operator=op3, Arrow=m3)

knitr::kable(t, caption="Custom operators in mathml",
  row.names=FALSE, escape=FALSE)
```

## Builtin functions

There is support for most functions from package _base_, with adequate use and
omission of parentheses.

```{r, results='asis'}
term <- quote(sin(x) + sin(x)^2L + cos(pi/2L) + tan(2L*pi) * expm1(x))
mathout(term)
term <- quote(choose(N, k) + abs(x) + sqrt(x) + floor(x) + exp(frac(x, y)))
mathout(term)
```

A few more examples are shown in Table \@ref(tab:base-stats), including
functions from `base` and `stats`.

```{r base-stats, echo=FALSE}
op1 <- list(
  "sin(x)"=quote(sin(x)),
  "cosh(x)"=quote(cosh(x)),
  "tanpi(alpha)"=quote(tanpi(alpha)),
  "asinh(x)"=quote(asinh(x)),
  "log(p)"=quote(log(p)),
  "log1p(x)"=quote(log1p(x)),
  "logb(x,\\ e)"=quote(logb(x, e)),
  "exp(x)"=quote(exp(x)),
  "expm1(x)"=quote(expm1(x)),
  "choose(n,\\ k)"=quote(choose(n, k)),
  "lchoose(n,\\ k)"=quote(lchoose(n, k)),
  "factorial(n)"=quote(factorial(n)), 
  "lfactorial(n)"=quote(lfactorial(n)),
  "sqrt(x)"=quote(sqrt(x)),
  "mean(X)"=quote(mean(X)),
  "abs(x)"=quote(abs(x)))
m1 <- lapply(op1, FUN=mathout, flags=list(cat=FALSE))

op1 <- names(op1)
if(knitr::is_latex_output())
  op1 <- sapply(op1, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op1 <- sapply(op1, FUN=knitr:::escape_html)

op2 <- list(
  "dbinom(k,\\ N,\\ pi)"=quote(dbinom(k, N, pi)),
  "pbinom(k,\\ N,\\ pi)"=quote(pbinom(k, N, pi)),
  "qbinom(p,\\ N,\\ pi)"=quote(qbinom(p, N, pi)),
  "dpois(k,\\ lambda)"=quote(dpois(k, lambda)),
  "ppois(k,\\ lambda)"=quote(ppois(k, lambda)),
  "qpois(p,\\ lambda)"=quote(qpois(p, lambda)),
  "dexp(x,\\ lambda)"=quote(dexp(x, lambda)),
  "pexp(x,\\ lambda)"=quote(pexp(x, lambda)),
  "qexp(p,\\ lambda)"=quote(qexp(p, lambda)),
  "dnorm(x,\\ mu,\\ sigma)"=quote(dnorm(x, mu, sigma)),
  "pnorm(x,\\ mu,\\ sigma)"=quote(pnorm(x, mu, sigma)),
  "qnorm(alpha/2L)"=quote(qnorm(alpha/2L)),
  "1L\\ -\\ pchisq(x,\\ 1L)"=quote(1L - pchisq(x, 1L)),
  "qchisq(1L\\ -\\ alpha,\\ 1L)"=quote(qchisq(1L-alpha, 1L)),
  "pt(t,\\ N\\ -\\ 1L)"=quote(pt(t, N-1L)),
  "qt(alpha/2L,\\ N\\ -\\ 1L)"=quote(qt(alpha/2L, N-1L)))
m2 <- lapply(op2, FUN=mathout, flags=list(cat=FALSE))

op2 <- names(op2)
if(knitr::is_latex_output())
  op2 <- sapply(op2, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op2 <- sapply(op2, FUN=knitr:::escape_html)

t <- cbind(Function=op1, Output=m1, Function=op2, Output=m2)
knitr::kable(t, caption="R functions from base and stats",
  row.names=FALSE, escape=FALSE)
```

## Custom functions

For self-written functions, matters are a bit more complicated. For a function
such as `g <- function(...) ...`, the name _g_ is not transparent to R, because
only the function body is represented, as illustrated by the redefinition
of `sign` below.

```{r, results='asis'}
sgn <- function(x)
{
  if(x == 0L) return(0L)
  if(x < 0L) return(-1L)
  if(x > 0L) return(1L)
}

mathout(sgn)
mathout(call("<-", quote(sgn(x)), sgn))
```

As shown in the example, we can still display functions in the
form `head(x) = body` if we embed the object to be shown into a
call `"<-"(head, body)`.

The function body is generally a nested R\ call of the form `'{'(L)`, with `L`
being a list of commands (the semicolon, not necessary in R, is translated to a
newline). As illustrated in the example, \pkg{mathml} provides limited support
for control structures such as `if`.

## Indices and powers

Indices in square brackets are rendered as subscripts, powers are rendered as
superscript. Moreover, \pkg{mathml} defines the
functions `sum_over(x, from, to)`, and `prod_over(x, from, to)` that simply
return their first argument. The other two arguments serve as
decorations (_to_ is optional), for example, for summation and product signs.

```{r, results='asis'}
term <- quote(S[Y]^2L <- frac(1L, N) * sum(Y[i] - mean(Y))^2L)
mathout(term)

term <- quote(log(prod_over(L[i], i==1L, N)) <- sum_over(log(L[i]), i==1L, N))
mathout(term)
```

## Ringing back to R

R\'s `integrate` function takes a number of arguments, the most important ones
being the function to integrate, and the lower and the upper bound of the
integration.

```{r, results='asis'}
term <- quote(integrate(sin, 0L, 2L*pi))
mathout(term)
eval(term)
```

For mathematical typesetting in the form
of \(\int f(x)\, dx\), \pkg{mathml} needs to find out the name of the
integration variable. For that purpose, the underlying Prolog bridge provides a
predicate `r_eval/3` that calls R\ from Prolog. In the example above, this
predicate evaluates `formalArgs(args(sin))`, which returns the names of the
arguments of `sin`, namely, `x`.

Note that in the example above, the quoted term is an abbreviation
for `call("integrate", quote(sin), ...)`, with `sin` being an R\ symbol, not a
function. While the R\ function `integrate()` can handle both symbols and
functions, \pkg{mathml} needs the symbol because it is unable to determine the
function name of custom functions.

## Names and order of arguments

One of R's great features is the possibility to refer to function arguments by
their names, not only by their position in the list of arguments. At the other
end, Prolog does not have such a feature. Therefore, the Prolog handlers for
R\ calls are rather rigid, for example, `integrate/3` accepts exactly three
arguments in a particular order and without names, that
is, `integrate(lower=0L, upper=2L*pi, sin)`, would not print the desired result.

To "canonicalize" function calls with named arguments and arguments in unusual
order, \pkg{mathml} provides an auxiliary R\ function `canonical(f, drop)` that
reorders the argument list of calls to known R\ functions and, 
if `drop=TRUE` (which is the default), also removes the names of the arguments.

```{r}
term <- quote(integrate(lower=0L, upper=2L*pi, sin))
canonical(term)
```

```{r, results='asis'}
mathout(canonical(term))
```

This function can be used to feed mixtures of partially named and positional
arguments into the renderer. For details, see the R\ function `match.call()`.

## Matrices and Vectors

Of course, \pkg{mathml} also supports matrices and vectors.

```{r, results='asis'}
v <- 1:3
mathout(call("t", v))

A <- matrix(data=11:16, nrow=2, ncol=3)
B <- matrix(data=21:26, nrow=2, ncol=3)
term <- call("+", A, B)
mathout(term)
```

Note that the seemingly more convenient `term <- quote(A + B)` yields \(A + B\)
in the output---instead of the desired matrix representation. This behavior is
expected because quotation of R calls also quote the components of the
call (here, _A_ and _B_).

## Short mathematical names for R symbols

In typical R\ functions, variable names are typically longer than just single
letters, which may yield unsatisfactory results in the mathematical output.

```{r, results='asis'}
term <- quote(pbinom(successes, Ntotal, prob))
mathout(term)

hook(successes, k)
hook(quote(Ntotal), quote(N), quote=FALSE)
hook(prob, pi)
mathout(term)
```

To improve the situation, \pkg{mathml} provides a simple hook that can be used
to replace elements (e.g., verbose variable names) of the code by concise
mathematical symbols, as illustrated in the example. To simplify notation,
the `quote` flag of `hook()` defaults to TRUE, and `hook()` uses non-standard
evaluation to unpack its arguments. If quote is FALSE, as shown above, the user
has to provide the quoted expressions. Care should be taken to
avoid recursive hooks such as `hook(s, s["A"])` that endlessly replace
the \(s\) from \(s_{\mathrm{A}}\) as
in \(s_{\mathrm{A}_{\mathrm{A}_{\mathrm{A}\cdots}}}\).

The hooks can also be used for more complex elements such as R\ calls, with
dotted symbols representing Prolog variables.

```{r, results='asis'}
hook(pbinom(.K, .N, .P), sum_over(dbinom(i, .N, .P), i=0L, .K))
mathout(term)
```

## Abbreviations

We consider the \(t\)-statistic for independent samples with equal variance. To
avoid clutter in the equation, the pooled variance \(s^2_{\mathrm{pool}}\) is
abbreviated, and a comment is given with the expression
for \(s^2_{\mathrm{pool}}\). For this purpose, \pkg{mathml} provides a
function `denote(abbr, expr, info)`, with `expr` actually being
evaluated, `abbr` being rendered, plus a comment of the
form "with `expr` denoting `info`".

```{r, results='asis'}
hook(m_A, mean(X)["A"]) ; hook(s2_A, s["A"]^2L) ;
hook(n_A, n["A"])
hook(m_B, mean(X)["B"]) ; hook(s2_B, s["B"]^2L)
hook(n_B, n["B"]) ; hook(s2_p, s["pool"]^2L)

term <- quote(t <- dfrac(m_A - m_B, 
    sqrt(denote(s2_p, frac((n_A - 1L)*s2_A + (n_B - 1L)*s2_B, n_A + n_B - 2L),
                "the pooled variance.") * (frac(1L, n_A) + frac(1L, n_B)))))
mathout(term)
```

The term is evaluated below. `print()` is needed because the return value of an
assignment of the form `t <- dfrac(...)` is not visible in R.

```{r}
m_A <- 1.5; s2_A <- 2.4^2; n_A <- 27; m_B <- 3.9; s2_B <- 2.8^2; n_B <- 20
print(eval(term))
```

## Context-dependent rendering

Consider an educational scenario in which we want to highlight a certain
element of a term, for example, that a student has forgotten to subtract the
null hypothesis in a \(t\)-ratio:

```{r, results='asis'}
t <- quote(dfrac(omit_right(mean(D) - mu[0L]), s / sqrt(N)))
mathout(t, flags=list(error="highlight"))
mathout(t, flags=list(error="fix"))
```

The R function `omit_right(a + b)` uses non-standard evaluation
techniques [e.g., @Wickham2019] to return only the left part an operation,
and cancels the right part. This may not always be desired, for example, when
illustrating how to fix the mistake.

For this purpose, the functions `mathml()` or `mathjax()` have an optional
argument `flags` which is a list with named elements. In this example, we use
this argument to tell \pkg{mathml} how to render such erroneous
expressions using the flag `error` which can be asis, highlight, fix, or ignore.
For more examples, see Table \@ref(tab:mistakes).

```{r mistakes, echo=FALSE}
op1 <- list(
  "omit_left(a\\ +\\ b)"=quote(omit_left(a + b)),
  "omit_right(a\\ +\\ b)"=quote(omit_right(a + b)),
  "list(quote(a),\\ quote(omit(b)))"=list(quote(a), quote(omit(b))),
  "add_left(a\\ +\\ b)"=quote(add_left(a + b)),
  "add_right(a\\ +\\ b)"=quote(add_right(a + b)),
  "list(quote(a),\\ quote(add(b)))"=list(quote(a), quote(add(b))),
  "instead(a,\\ b)\\ +\\ c"=quote(instead(a, b) + c))

asis <- lapply(op1, FUN=mathout, flags=list(cat=FALSE, error="asis"))
high <- lapply(op1, FUN=mathout, flags=list(cat=FALSE, error="highlight"))
fix  <- lapply(op1, FUN=mathout, flags=list(cat=FALSE, error="fix"))
igno <- lapply(op1, FUN=mathout, flags=list(cat=FALSE, error="ignore"))

op1 <- names(op1)
if(knitr::is_latex_output())
  op1 <- sapply(op1, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op1 <- sapply(op1, FUN=knitr:::escape_html)

t <- cbind(Operation=op1, 
  "error\\ =\\ asis"=asis, highlight=high, fix=fix, ignore=igno)
knitr::kable(t, caption="Highlighting elements of a term",
  row.names=FALSE, escape=FALSE)
```

## Customizing the package

In typical R\ functions, variable names are typically longer than just single
letters, which may yield unsatisfactory results in the mathematical output.

```{r, results='asis'}
hook(successes, k)
hook(quote(Ntotal), quote(N), quote=FALSE)
hook(prob, pi)

term <- quote(dbinom(successes, Ntotal, prob))
mathout(term)
```

To improve the situation, \pkg{mathml} provides a simple hook that can be used
to replace elements (e.g., verbose variable names) of the code by concise
mathematical symbols, as illustrated in the example. To simplify notation,
the `quote` flag of `hook()` defaults to TRUE, and `hook()` uses non-standard
evaluation to unpack its arguments. If quote is FALSE, as shown above, the user
has to provide the quoted expressions. Care should be taken to
avoid recursive hooks such as `hook(s, s["A"])` that endlessly replace
the \(s\) from \(s_{\mathrm{A}}\) as
in \(s_{\mathrm{A}_{\mathrm{A}_{\mathrm{A}\cdots}}}\).

Further customization requires the assertion of new Prolog rules `math/2`,
`ml/3`, `jax/3`, as shown in the Appendix.

# A case study {#case-study}

This case study describes a model by Schwarz [-@schwarz1994] from mathematical
psychology using the features of package \pkg{mathml}. Schwarz presents a new
explanation of redundancy gains that occur when observers respond to stimuli of
different sources, and the same information is presented on two or more
channels. In Schwarz' model, decision-making builds on a process of noisy
accumulation of information over time [e.g., @ratcliff2016]. In redundant
stimuli, the model assumes a superposition of channel-specific diffusion
processes that eventually reach an absorbing barrier to elicit the response. For
a detailed description the reader may refer to the original article.

Schwarz' [-@schwarz1994] model refers to two stimuli A and B, presented either
alone or in combination (AB, redundant stimuli), with the redundant stimuli
being presented either simultaneously or with onset asynchrony \(\tau\). The
channel activation is described as a two-dimensional Wiener process with
drifts \(\mu_i\), variances \(\sigma^2_i\), and initial
conditions \(X_i(t = 0) = 0, i = \mathrm{A, B}\). The buildup of 
channel-specific activation may be correlated with \(\rho_{\mathrm{AB}}\), but
we assume \(\rho_{\mathrm{AB}} = 0\) for simplicity.

A response is elicited when the process reaches an absorbing barrier \(c > 0\)
for the first time. In single-target trials, the first passages of \(c\) are
expected at

```{r, results='asis'}
ED_single <- function(c, mu)
  dfrac(c, mu)

# display as E(D; mu), c is a scaling parameter
hook(ED_single(.C, .Mu), E(`;`(D, .Mu)))
mathout(call("=", quote(ED_single(c, mu)), ED_single))
```

(One would typically use chunk option `echo=FALSE` to suppress the R code.)
In redundant stimuli, the activations from the channel-specific diffusion
processes add up, \(X_{\mathrm{AB}}(t) = X_{\mathrm A}(t) + X_{\mathrm B}(t)\),
hence the name, superposition. \(X_{\mathrm{AB}}(t)\) is again a Wiener process
with drift \(\mu_{\mathrm A} + \mu_{\mathrm B}\) and
variance \(\sigma^2_{\mathrm A} + \sigma^2_{\mathrm B}\). For the expected
first-passage time, we have

```{r, results='asis'}
hook(mu_A, mu["A"])
hook(mu_B, mu["B"])
hook(sigma_A, sigma["A"])
hook(sigma_B, sigma["B"])
hook(mu_M, mu["M"])
hook(M, overline(X))

mathout(call("=", quote(E(D["AB"])), quote(ED_single(c, mu_A + mu_B))))
```

For asynchronous stimuli, Schwarz [-@schwarz1994] derived the expected
first-passage time as a function of the stimulus onset asynchrony \(\tau\),

```{r, results="asis"}
ED_async <- function(tau, c, mu_A, sigma_A, mu_B)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) *
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*c*mu_A, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

hook(ED_async(.Tau, .C, .MA, .SA, .MB), E(`;`(D[.Tau], `,`(.MA, .SA, .MB))))
mathout(call("=", quote(ED_async(tau, c, mu_A, sigma_A, mu_B)), ED_async))
```

For negative onset asynchrony (i.e., B before A), the parameters are simply
switched.

```{r, results='asis'}
ED <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{
  if(tau == Inf) return(ED_single(c, mu_A))
  if(tau == -Inf) return(ED_single(c, mu_B))
  if(tau == 0L) return(ED_single(c, mu_A + mu_B))
  if(tau > 0L) return(ED_async(tau, c, mu_A, sigma_A, mu_B))
  if(tau < 0L) return(ED_async(abs(tau), c, mu_B, sigma_B, mu_A))
}

hook(ED(.Tau, .C, .MA, .SA, .MB, .SB), E(`;`(D[.Tau], `,`(.MA, .SA, .MB, .SB))))
mathout(call("=", quote(ED(tau, c, mu_A, sigma_A, mu_B, sigma_B)), ED))
```

The observable response time is assumed to be the sum of \(D\), the time
employed to reach the threshold for the decision, and a residual \(M\) denoting
other processes such a motor preparation and execution. Correspondingly, the
expected response time amounts to

```{r, results='asis'}
ET <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
  ED(tau, c, mu_A, sigma_A, mu_B, sigma_B) + mu_M

hook(ET(.Tau, .C, .MA, .SA, .MB, .SB, .MM),
     E(`;`(T[.Tau], `,`(.MA, .SA, .MB, .SB, .MM))))
mathout(call("=", quote(E(T[tau])), ET))
```

Schwarz [-@schwarz1994] applied the model to data from a redundant signals task
with 13 onset
asynchronies \(0, \pm33, \pm67, \pm100, \pm133, \pm167, \pm\infty\) ms,
where \(\tau = 0\) refers to the synchronous condition, and \(\pm\infty\) to the
single-target presentations. Each condition was replicated 400 times. The
observed mean response times and their standard deviations are given in
Table \@ref(tab:miller-data) below.

```{r miller-data, echo=FALSE}
tau  <- list(-Inf,-167L,-133L,-100L,-67L,-33L,0L,33L,67L,100L,133L,167L,Inf)
m    <- c(231, 234, 230, 227, 228, 221, 217, 238, 263, 277, 298, 316, 348)
s    <- c( 56,  58,  40,  40,  32,  28,  28,  28,  26,  30,  32,  34,  92)
n    <- c(400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400)

ttau <- lapply(tau, FUN=mathout, flags=list(cat=FALSE))
tm   <- lapply(as.integer(m), FUN=mathout, flags=list(cat=FALSE))
ts   <- lapply(as.integer(s), FUN=mathout, flags=list(cat=FALSE))
tn   <- lapply(as.integer(n), FUN=mathout, flags=list(cat=FALSE))

t <- cbind('\\(\\tau\\)'=ttau, '\\(m\\)'=tm, '\\(s\\)'=ts, '\\(n\\)'=tn)
knitr::kable(t, caption="Miller data", row.names=FALSE, escape=FALSE)
```

Assuming that the model is correct, the observable mean reaction times follow an
approximate Normal distribution around the model prediction \(E(T_\tau)\) for
each condition. We can, therefore, use a standard goodness-of-fit measure
by \(z\)-standardization.

```{r, results='asis'}
z <- function(m, s, n, tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
  dfrac(m - denote(mu[tau], ET(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M),
              "the expected mean response time"),
    s / sqrt(n))

mathout(call("=", quote(z[tau]), z))
```

The overall goodness-of-fit is the sum of the squared \(z\)-statistics for each
onset asynchrony. Assuming again that the architecture of the model is correct,
but the parameters are adjusted to the data, it follows
a \(\chi^2(8\ \mathrm{df})\)-distribution.

```{r, results='asis'}
zv <- Vectorize(z, vectorize.args = c('m', 's', 'n', 'tau'))
hook(zv(.M, .S, .N, .Tau, .C, .MA, .SA, .MB, .SB, .MM), z[.Tau])

gof <- function(par, tau, m, s, n)
  sum(zv(m, s, n, tau, c=100L, mu_A=par["mu_A"], sigma_A=par["sigma_A"], 
         mu_B=par["mu_B"], sigma_B=par["sigma_B"], mu_M=par["mu_M"])^2L)

mathout(call("=", quote(X["8 df"]^2L), gof))
```

with the degrees of freedom given by the difference between the number of
observations (`r length(tau)`) and the number of free model
parameters \(\theta = \langle\mu_{\mathrm A}, \sigma_{\mathrm A},
\mu_{\mathrm B}, \sigma_{\mathrm B}, \mu_{\mathrm M}\rangle\); the barrier \(c\)
is only a scaling parameter.

```{r, results="asis", echo=FALSE}
lsq <- function(tau, m, s, n)
{
  invisible(theta <- c(mu_A=0.5, sigma_A=5, mu_B=0.5, sigma_B=5, mu_M=150))
  optim(theta, gof, tau=tau, m=m, s=s, n=n, hessian=TRUE)
}

mathout(call("=", quote(hat(theta)), lsq))
```

The best fitting parameter values and their confidence intervals are given in
Table \@ref(tab:params) below.

```{r params, echo=FALSE}
fit <- lsq(tau, m, s, n)
theta <- fit$par
se <- sqrt(diag(solve(fit$hessian)))
ci <- mapply(FUN=qnorm, theta, se, MoreArgs=list(p=c(0.025, 0.975)),
    SIMPLIFY=FALSE)

ntheta <- lapply(lapply(FUN=as.symbol, names(theta)), FUN=mathout,
    flags=list(cat=FALSE))
ttheta <- lapply(theta, FUN=mathout, flags=list(cat=FALSE))
tci <- lapply(ci, FUN=mathout, flags=list(cat=FALSE, sep=","))

t <- cbind(Parameter=ntheta, Estimate=ttheta, CI=tci)
knitr::kable(t, caption="Model fit", row.names=FALSE, escape=FALSE)
```

The goodness-of-fit statistic indicates some lack of
fit, \(X^2(8\ \mathrm{df}) = `r sprintf("%.2f", fit$value)`, 
  p = `r sprintf("%.4f", 1-pchisq(fit$value, df=8))`\). Given the large trial
numbers in the original study, this is not an unexpected result. For more
detail, especially on fitting the observed standard deviations, the reader is
referred to the original paper [@schwarz1994].

# Conclusions

This package allows R\ to render its terms in pretty mathematical equations. It
extends the current features of R\ and existing packages for displaying
mathematical formulas in R\ [@murrell2000;@allaire2018]), bit most
importantly, \pkg{mathml} bridges the gap between computational needs, presentation
of results, and their reproducibility. The package supports both MathML and
Latex/MathJax for use in R\ Markdown documents, presentations and ShinyApp
webpages.

Researchers or teachers can already use RMarkdown to conduct analyses
and show results, and \pkg{mathml} smoothes this process and allows for integrated
calculations and output. As shown in the case study of the previous
section, \pkg{mathml} can help to improve data analyses and statistical reports from
an aesthetical perspective, as well as regarding reproducibility of research.

Furthermore, the package may also allow for a better detection of possible
mistakes in R\ programs. Similar to most programming languages [@green1977],
R\ code is notoriously hard to read, and the poor legibility of the language is
one of the main sources of mistakes. For illustration, we consider again
Equation\ 10 in Schwarz [-@schwarz1994].

````{r, results="asis"}
f1 <- function(tau)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B) * 
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*tau, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau)))))
}

mathout(f1)
````

The first version has a wrong parenthesis, which is barely visible in the code,
whereas in the mathematical representation, the wrong curly brace is immediately
obvious (the correct version is shown below for comparison).

````{r, results="asis"}
f2 <- function(tau)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) * 
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*tau, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

mathout(f2)
````

As the reader may know from own experience, missed parentheses are frequent
causes of wrong results and errors that are hard to locate in programming code.
This particular example shows that mathematical rendering can help to
substantially reduce the amount of careless errors in programming.

In its current version \pkg{mathml} has some limitations. For example, it is
currently not possible to use the functions of \pkg{mathml} for writing
inline formulas, here, the user has to adopt the usual LateX notation.
Moreover, for long equations, we did not yet find a convenient way to insert
line breaks. This is mostly due to lacking support by MathML and LaTeX
renderers. For example, in its current stage, the LaTeX package `breqn` [@breqn]
is mostly a proof of concept.

The package \pkg{mathml} is available for R\ version (…) and later, and can be
easily installed using the usual `install.packages("mathml")`. The source code
of the package is found at https://github.com/mgondan/mathml.

# Appendix: Customizing the package

Several ways exist for translating new R\ terms to their mathematical
representation. We have already seen above how to use "hooks" to translate long
variable names from R to compact mathematical signs, as well as functions such
as cumulative probabilities \(P(X \le k)\) to different representations
like \(\sum_{i=0}^k P(X = i)\). Obviously, the hooks require that there already
exists a rule to translate the target representation into MathML and MathJax.

In this appendix we describe a few more ways to extend the set of translations
according to a user's needs. As stated in the background section, the Prolog end
provides two classes of rules for translation, macros `math/2,3,4` mirroring the
R\ hooks mentioned above, and the low-level predicates `ml/3` and `jax/3` that
create proper MathML and Latex terms.

## Linear models

To render the model equation of a linear model such
as `lm(EOT ~ T0 + Therapy, data=d)` in mathematical form, it is sufficient to
map the `Formula` in `lm(Formula, Data)` to its respective equation. This can in
two ways, using either the hooks described above, or a new `math/2` macro at the
Prolog end.

````r
hook(lm(.Formula, .Data), .Formula)
````

The hook is simple, but is a bit limited because only R's tilde-form
of linear models is shown, and it only works for a call with exactly two
arguments.

Below is an example how to build a linear equation of the
form \(Y = b_0 + b_1X_1 + ...\) using the Prolog macros from \pkg{mathml}.

````prolog
math_hook(LM, M) :-
    compound(LM),
    LM =.. [lm, ~(Y, Sum) | _Tail],
    summands(Sum, Predictors),
    findall(subscript(b, X) * X, member(X, Predictors), Terms),
    summands(Model, Terms),
    M = (Y == subscript(b, 0) + Model + epsilon).
````

The predicate `summands/2` unpacks an expression `A + B + C` to a
list `[C, B, A]` and vice-versa (see the file `lm.pl` for details).

````{r, results='asis'}
rolog::consult(system.file(file.path("pl", "lm.pl"), package="mathml"))

term <- quote(lm(EOT ~ T0 + Therapy, data=d, na.action=na.fail))
mathout(term)
````

## \(n\)-th root

Base R does not provide a function like `cuberoot(x)` or `nthroot(x, n)`, and
the present package does not support the respective representation. To obtain a
cube root, a programmer would typically type `x^(1/3)` or better `x^{1/3}` (see
the practice section why the curly brace is preferred in an exponent), resulting
in \(x^{1/3}\) which may still not match everyone's taste. Here describe the
steps needed to represent the \(n\)-th root as \(\sqrt[n]x\).

We assume that `nthroot(x, n)` is available in the current
namespace [manually defined, or from R package \CRANpkg{pracma}, @pracma], so that
the names of the arguments and their order are accessible to `canonical()` if
needed. As we can see below, \pkg{mathml} uses a default
representation `name(arguments)` for such unknown functions.

```{r, results='asis'}
nthroot <- function(x, n)
  x^{1L/n}

term <- canonical(quote(nthroot(n=3L, 2L)))
mathout(term)
```

A proper MathML term is obtained by `mlx/3` (the x in mlx indicates that it is
an extension and is prioritized over the default ml/3 rules). `mlx/3`
recursively invokes `ml/3` for translating the function arguments _X_ and _N_,
and then constructs the correct MathML entity `<mroot>...</mroot>`.

```prolog
mlx(nthroot(X, N), M, Flags) :-
    ml(X, X1, Flags),
    ml(N, N1, Flags),
    M = mroot([X1, N1]).
```

The explicit unification `M = ...` in the last line serves to avoid clutter in
the head of `mlx/3`. The Prolog file `nthroot.pl` also includes the respective
rule for Latex and can be consulted from the package folder via the underlying
package \pkg{rolog}.

```{r, results='asis'}
rolog::consult(system.file(file.path("pl", "nthroot.pl"), package="mathml"))

term <- quote(nthroot(a * (b + c), 3L)^2L)
mathout(term)
term <- quote(a^(1L/3L) + a^{1L/3L} + a^(1.0/3L))
mathout(term)
```

The file `nthroot.pl` includes three more statements `precx/3` and `parenx/3`,
as well as a `math_hook/2` macro. The first sets the operator precedence of the
cubic root above the power, thereby putting a parentheses around nthroot
in \((\sqrt[3]{\ldots})^2\). The second tells the system to increase the counter
of the parentheses below the root, such that the outer parenthesis becomes a
square bracket.

The last rule maps powers like `a^(1L/3L)` to `nthroot/3`, as shown in the
first summand. Of course, \pkg{mathml} is not a proper computer algebra system.
As is illustrated by the other terms in the sum, such macros are limited to
purely syntactical matching, and terms like `a^{1L/3L}` with the curly brace
or `a^(1.0/3L)` with a floating point number in the numerator are not detected.

# Acknowledgment

Supported by the Erasmus+ program of the European
Commission (2019-1-EE01-KA203-051708).